C/O and Snowline Locations in Protoplanetary Disks: The Effect of Radial Drift and Viscous Gas Accretion

### 2015-11-14 tex/draft/ms_final_revised_2.tex
update

\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis. For the evolving and viscous disks, the C/O ratio in gas may decrease due to the decline in the surface density of solids with time. Thus both the relative fluxes of the volatiles and the decrease in solid ice abundances will lower the C/O ratio in evolving and viscous disks, as indicated by the arrows in Figure \ref{fig:CO_ratio}. However, a high gas-phase C/O ratio is maintained further in for all disks compared to a static disk, since drift and gas accretion cause the ices to desorb closer to the host star.}

@@

>>>>

%\textbf{Our premise that each volatile has a fixed abundance at every location in the disk yields good approximations for the C/O ratio in an irradiated disk --- our model assumes a constant influx of particles at any given radius, implying that the solid surface density of a given volatile, and therefore the amount of desorbed vapor, remain constant with semimajor axis. For the evolving and viscous disks, the C/O ratio in gas may decrease due to the decline in the surface density of solids with time. Thus both the relative fluxes of the volatiles and the decrease in solid ice abundances will lower the C/O ratio in evolving and viscous disks, as indicated by the arrows in Figure \ref{fig:CO_ratio}. However, a high gas-phase C/O ratio is maintained further in for all disks compared to a static disk, since drift and gas accretion cause the ices to desorb closer to the host star.}

%\textbf{Desorption is taken into account in the figure, and fundamentally the elevated C/O ratio interior to the static CO2 and H2O snowlines are simply due to the inward mobility of the snowlines (desorption fronts) due to drift and accretion flows. Qualitatively this scenario should be robust to changes in total abundances throughout the disk, i.e. at e.g. the ‘dynamic’ CO2 snowlines, the rapid return of CO2 into the gas-phase (during CO2 desorption) will reduce the C/O ratio interior to the CO2 dynamic snowline, while no major change in gas-phase composition and therefore C/O ratio is expected between the static and dynamic snowlines.

%As the referee notes, we do operate under the simplifying assumption that the total (ice+gas) abundances are the same at each radius after ices have migrated. This is a good
%approximation for the irradiated disk, given that this model by definition presents a constant influx of particles at any given radius while the gas is static, and thus the ice+gas surface density should remain constant. For the evolving disk, we agree with the referee that this is not a good approximation. In evolving disks, the gas-phase C/O ratio may decrease everywhere interior to the CO2 and H2O desorption fronts due to the decrease in the surface density of solids with time at any given radius. In the steady-state viscous disk, the solid abundances at a fixed radius are constant, given that this model is
%not time-dependent, but the solid/gas ratio is not constant, which can result in a substantially lower C/O ratio interior to the H2O and CO2 snowlines compared to the static case (as indicated by the arrows in the figure).

%These caveats are important and we have explained and clarified them further in the next. We have also clarified that the main purpose of this figure is to show the different snowline locations in static and dynamic disks, and thus where in the disk C/O is reduced or increased rather than providing a quantitative estimate of how big or small that increase/decrease is.}

@@

\textbf{Fundamentally, the elevated C/O ratios interior to the static H$_2$O and CO$_2$ snowlines are simply caused by the inward movement of the snowlines due to radial drift and gas accretion. Qualitatively, this scenario should be robust to changes in total abundances throughout the disk --- for example, at the dynamic (non-static) CO$_2$ snowlines, the rapid return of CO$_2$ into gas-phase during CO$_2$ desorption will reduce the C/O ratio interior to the CO$_2$ dynamic snowline, while no major change in gas-phase composition, and therefore C/O ratio, is expected between the static and dynamic snowlines.}

\textbf{As noted earlier, we assume that the total (ice and gas) abundance of each volatile is the same at every radius after ices have migrated. This is a good approximation for the irradiated disk, given that this model, by definition, presents a constant influx of particles at any given radius while the gas is static, and thus the ice and gas surface density remain constant. For the evolving disk, the gas-phase C/O ratio may decrease everywhere interior to the H$_2$O and CO$_2$ snowlines due to the decline in the surface density of solids with time at any given radius. For the viscous disk, the solid abundances at a fixed radius are constant, given that this model is not time-dependent, but the gas-to-solid ratio is not constant, which can result in a substantially lower C/O ratio interior to the H$_2$O and CO$_2$ snowlines compared to the static case (as indicated by the arrows in Figure \ref{fig:CO_ratio}). The main goal of Figure \ref{fig:CO_ratio}, however, is to show the different snowline radii in static and non-static disks, and therefore the locations in the disk where the gas-phase C/O ratio is reduced or increased, rather than provide a quantitative estimate of the magnitude of the C/O increase or decrease.}
